Paper analysis

### 2025-03-28 Paper_analysis.Rmd
update site

# Interactions
```{r int_load}
library(iml)
library(gridExtra)
load("data/ES-RF2-preds.RDA")
```

## Pollinator base interactions

```{r polbase_int}
grid.arrange(
  int.polbase %>% plot() + ggtitle("Interaction strength") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  int.polbase.ntr_pland %>% plot() + ggtitle("Pairwise interactions with strongest") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  ncol = 2
)
```

## Pollinators margins interactions

```{r polmarg_int}
grid.arrange(
  int.polmarg %>% plot() + ggtitle("Interaction strength") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  int.polmarg.ntr_pland %>% plot() + ggtitle("Pairwise interactions with strongest") + egg::theme_article() + theme(plot.title.position = "plot", axis.title.y = element_blank()),
  ncol = 2
)
```

>>>>

```{r impurity_importance, fig.height=16, fig.width=9}

load("data/ES-RF2-impurity.RDA")

library(ggpattern)

imp.imp.df <- bind_rows(
  pollinators_base.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Pollinators", model = "base"),
  pollinators_margin.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Pollinators", model = "margin"),
  n_export_base.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "N export", model = "base"),
  n_export_margin.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "N export", model = "margin"),
  sed_export_base.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Sed export", model = "base"),
  sed_export_margin.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Sed export", model = "margin"),
  qb_base.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Baseflow", model = "base"),
  qb_margin.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Baseflow", model = "margin")
) %>%
  mutate(sig = ifelse(pvalue <0.05, TRUE, FALSE)) %>%
  mutate(metric2 = tidytext::reorder_within(metric, importance,
                                           within = list(ES, model))) %>%
  left_join(metric.descr %>% select(lsm, type), by = c("metric"="lsm")) %>%
  mutate(type = case_match(
    metric,
    "field_pland" ~ "area and edge metric",
    "field_frac_mn" ~ "shape metric",
    "field_frac_cv" ~ "shape metric",
    "prn_ann_ratio" ~ "area and edge metric",
    .default = type
  ))

ggplot(imp.imp.df, aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  ggh4x::facet_grid2(ES~model, scales = "free", independent = "all") +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

## Shapley importance
```{r importance_shap}
load("data/ES-RF2-shapimportance.RDA")
library(shapviz)
library(gridExtra)

### Pollinators

*Pollinators*

```{r shap_pol}
grid.arrange(
  sv_importance(ex.polbase.glb), sv_importance(ex.polmarg.glb),
  ncol = 2
)

*N Export*

```{r shap_Nexp}
grid.arrange(
  sv_importance(ex.nexpbase.glb), sv_importance(ex.nexpmarg.glb),
  ncol = 2
)

*Sediment Export*

```{r shap_sed}
grid.arrange(
  sv_importance(ex.sedbase.glb), sv_importance(ex.sedmarg.glb),
  ncol = 2
)

*Water recharge*

```{r shap_wb}
grid.arrange(
  sv_importance(ex.qbbase.glb), sv_importance(ex.qbmarg.glb),
  ncol = 2
)
